---
title: "Alewife Migration"
authors: "Danni Harper, Jillian Hunt, Jessie Lilly"
date: "February 12, 2018"
output: html_document



---
```{r}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<<<<<<< HEAD

# **Background:** 

Alewife, *Alosa pseudoharengus* (A. Wilson, 1811) is an anadromous species that is found along the Eastern coast of North America, migrating as far north as Newfoundland (Bigelow and Schroeder 1953). Alewife live to 10 years of age and begin spawning at the age of 3. Alewife can spawn up to 5 times throughout its life. Spawning begins in April or May, with the run reaching a peak in early June (DFO 2007). Alewife spawn upriver in stillwater, spending 13-14 days in the spawning ground before returning to sea (DFO,2007; Andrews 2014). 

The data for our study were collected from Amherst fishway and Missaquash fishway between April- June (2013-2017). The Amherst fishway is located just outside of Amherst, Nova Scotia and the Missaquash fishway is located on the border between Nova Scotia and New Brunswick, both fishways are Denil style fishways (Andrews 2014). For this study, Alewives were obtained using fyke nets and were transferred to a holding pen where measurements were taken, which included: fork length, total length, weight and sex. A small incision was made anterior to the anal fin and Pit tags were inserted into the abdominal cavity(Andrews 2014). Reader boxes were installed directly below each fishway to monitor PIT tagged Alewife passage frequency during their spawning run.

In a previous study it was noted that the passage of an Alewife over Missasquash and Ahmerst fish ways was positively correlated with the size of the fish (Andrews 2014). Larger fish were also more likely to overcome the barriers than smaller fish (Andrews 2014). It was also noted that Alewives often struggled to make it past the fish passages on their first attempt (Andrews 2014). This is hypothesized to occur because some fish do not have enough stored energy to overcome the barrier (Task and Fisher 1977). This suggests that these fishways may prevent Alewife from having successful spawning runs in the Ahmerst and Missaquash rivers. 



# **References**

Andrews, S.N. 2014. Fishway Efficiency and Passage Behavior of Alewife in Three        Fishways on Tantramar Marsh Near Amherst, Nova Scotia. Thesis.
Bigelow, H.., and Schroeder, W.C. 1953. Fishes of the Gulf of Maine. Fish. Bull. Fish  Wildl. Serv. 53. Available from http://www.gma.org/fogm/Microgadus_tomcod.htm.
DFO. (2007). Assessment of Gasperaue River Alewife. DFO Can. Sci. Advis. Sec. Sci.     Advis. Rep. 2007/030.
Task, S.., and Fisher, J.. 1977. Performance of Arctic Grayling in a twenty-foot       section model “A” Alaska Steeppass fish ladder. Final Rep. US Army Corps Eng. Alsk.    Div. Anchorage.




# **Questions**

Since it appears that larger fish are more successful than smaller fish at completing migrations over fishways, the objective of this analysis will be to: 

<<<<<<< HEAD
1. JILL Is there a relationship between forklength of fish and sex? --> have to do a t-test
  Response variable -> forklength (normally distributed)
  Predictor variable -> sex

2. Is there a relationship between sex, migration time?
  Response variable -> migration time
  Predictor variable -> sex
  
  2b) Is there a relationship between sex and size class?
    Response variable -> size class
    Predictor variable -> sex
  
3. DANNI Is there a difference between size of fish and tagging location?
  Response variable -> tagging location
  Predictor variable -> size of fish 
  
4. JESSIE Is there a difference between migration time and amount of fish (per day, per location)? -> Normal linear regression
Is there a difference between size of fork length and month?
=======
1. Is there a relationship between forklength of fish and sex? 
2. Is there a relationship between sex, migration time and size class?
3. Is there a difference between size of fish and tagging location?
>>>>>>> 9f8b9d8cb83d5c51a47e3732bb8a5572e0cf4ba2



# **Variables:** 
fl= forklength
tl=taillength
sx=sex
tl/fl= tail length to fork length ratio
w=weight
p= pit tag
h= hour
min= minute
d=day
m= month
y=year 
site= 
cblc: Ahmerst river tidal gate
ms: Missassguash tidal gate 

# **Data Tidying**
# We initially checked the status of each column and noted that pit was read in as an integer.


parsing failure: pit = double; sx = integer; sp = integer; site = integer...WHY? Check the status of each column. 

```{r}
library("tidyverse")

alewife17 <- read.csv("alewife13_17.csv", na = c("na", "NA"))
names(alewife17)

sapply(alewife17,class)



```

##We then changed the pit column from numeric to a character vector

```{r}

alewife17$pit <- as.character(alewife17$pit)
```

##Filter only 2017 data 

```{r}
alewife17.df <- filter(alewife17, y==2017)
```



Fill in missing values


```{r}

alewife17.df %>%
  count(sx)

```

```{r}
aw.df <- mutate(alewife17.df, 
                sex = toupper(str_trim(sx)), 
                sex = ifelse(sex %in% c("-","?", "NA"), "U", sex))
# check
with(aw.df, table(sex, sx))
```

Remove sx column from aw.df

```{r}
aw.df$sx <- NULL
```

Change na to a missing value in w

```{r}
library(tidyverse)
aw.df <- mutate(aw.df,w = ifelse(w %in% c("na"), "NA", w))
View(aw.df)
```

Below is a scatterplot displaying the relationship between forklength versus weight of Alewives grouped by sex. I am not sure why the numbers on the x axis are not displaying correctly. 
 
```{r}
ggplot(data= aw.df,mapping= aes(x=w,y=fl,colour=sex),position="jitter")+
  geom_point()+
  ggtitle("Fork length versus Weight for Alewives")+
  xlab("Weight(g)")+
  ylab("Fork length(fl)")

```

```{r}
library(lubridate)

aw.df %>%
  mutate(month=month(m,label=TRUE)) %>%
  ggplot(aes(x=m))+
  geom_bar()+
  ggtitle("Month vs. Passage of Alewife Through Ahmerst and Missaquash Fishway")+
  xlab("Month")+
  ylab("Count")
```

Graph showing the number of fish entering the system by date, and bars are separated by sex
```{r}

aw.df <- aw.df %>%
  mutate(enter = make_date(y, m, d))

ggplot(aw.df, aes(enter))+
  geom_histogram(aes(fill = sex), binwidth = 1)+
  xlab("Date")+
  ylab("Number of Fish")

```
```{r}
aw.df <- aw.df %>%
  mutate(length = ntile(fl, 6))

View(aw.df)
```

Graph showing the size of fish entering the system, colour is just the binned forklength (kinda redundant...)
```{r}
ggplot(aw.df, aes(enter, fl, alpha = 1/10))+
  geom_point()+
  ylab("Forklength (mm)")+
  xlab("Entery Date")+
  theme(legend.position = 'none')
```


## Change NA values to U

```{r}
aw.df.2 <- mutate(aw.df,sex = ifelse(sex %in% c("NA"),"U", sex))
```




## Graph of the relationship between forklength and sex. 

```{r}



ggplot(data=aw.df.2,mapping=aes(x=sex,y=fl))+
  geom_boxplot()+
  ggtitle("Relationship between Forklength and Sex")+
  xlab("Sex")+
  ylab("Forklength")


```

##Graph the relationship between size (fl) and tagging location (site)
```{r}
ggplot(data = aw.df) +
  geom_boxplot(mapping = aes(x = site, y = fl)) +
  ggtitle("Relationship between Fork Length and Tagging Location") + 
  ylab("Fork Length") +
  xlab("Tagging Location")
```

OUR RESEARCH QUESTIONS 

1. Is there a relationship between forklength of fish and sex?
  Response variable -> forklength (normally distributed)
  Predictor variable -> sex
  
  Using a Welsh two-sample t-test -> I know this isn't the ideal test to perform, but it will determine if the two groups (male and female) have the same mean of forklength. If the t-test indicates that the mean is not the same, we can determine that the forklengths differ among sex
  
Create two datasets from aw.df; one with just female forklenghs and the other with just male forklengths
```{r}
newmodel.df <- select(aw.df, fl, sex)
female.df <- filter(newmodel.df, sex == "F")
male.df <- filter(newmodel.df, sex == "M")
female.df$sex = NULL
male.df$sex = NULL
colnames(female.df)[1] = "female.fl"
colnames(male.df)[1] = "male.fl"
```

We now want to create a new dataset by combining the forklength columns of male.df and female.df, but there's an issue: male.df has 763 values and female.df has 730 values. So we first need to add 33 values of NA to the female dataset -- I know there's a "one-line code" way to do this but I couldn't figure it out, so I did 33 seperate lines of code to add them all individually.

```{r}
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
female.df <- rbind(female.df, "731" = c(NA, NA))
```

Create the new dataset with the columns male.fl and female.fl

```{r}
male.df$female.fl <- female.df$female.fl
female.fl <- male.df$female.fl
male.fl <- male.df$male.fl
```

Now we can run the t.test

```{r}
t.test(female.fl, male.fl)
```


3.Is there a difference between size of fish and tagging location?
  Response variable -> forklength of fish
  Predictor variable -> site
 
Creating a data frame with just the information I will need for the model (not necessary but easier for me to see)
```{r}
aw.sites <- select(aw.df, fl, site)
```

Test to see if ANOVA of this data frame works
```{r}
aov(fl ~ site, aw.sites)
```

Creating linear model to look at relationship between forklength (fl) and site 
```{r}
mod.site <- lm(fl ~ site, aw.sites)
```

Doing ANOVA on the model created above to see whether forklength is different by location
```{r}
anova(mod.site)
```
Based on the ANOVA there is not evidence to suggest that fish of different size prefer different fishways

Summary of my model
```{r}
summary(mod.site)
```

Residual plots of this model
```{r}
par(mfrow = c(2, 2))
plot(mod.site)
```
Model seems to fit pretty well, the values are evenly distributed around zero and Q-Q plot follows the line with a bit of spread at the top and bottom

=======
## 4. JESSIE Is there a difference between migration time and amount of fish observed?

## I think we should use a poisson distribution as we have different times of day in which fish were captured and a posssion test is used for count data. We want to compare whether the abundance of Alewife differs between different hours of the day, and this measure is a discrete variable. The fish are assumed to be distibuted randomly and independently and samples of equal size were assumed to be taken randomly. 

## Create a histogram to compare abundance. 

```{r}
ggplot(data=aw.df,aes(h))+
  geom_histogram()


```


## Count the Alewife Present at each hour
## The data appears to be skewed to the right. 

```{r}
data.frame(table(aw.df$h))
```

## Fit the model 
## I created a data frame that contains the frequncy of alweife captured at each hour. The frequency needs to be used in the poisson glm. 

```{r}
aw.df.2 <- aw.df %>% group_by(h) %>%tally()
```

## Use poisson distribution.
## I am not sure how to interpret the coefficients.  
## There will be a 3.1% increase in the number of fish caught for each increase in hour throughout the day? 
## There appears to be a statistically significant difference between different hours of the day and the number of Alewife captured? 
.

```{r}

p1 <- glm(n~h,family="poisson",data=aw.df.2)
summary(p1)
```

## Check assumptions
## Residuals versus fitted values: The indendence of error assumption can't be followed, as some values appear to influence one another. Do you need this in a poission model because the variance is equal to the mean? 
## The residuals are normally distributed. 
. 

```{r}
par(mfrow=c(2,2))
plot(p1)
```
