---
title: "Fisheries_class"
author: "Jessie Lilly"
date: "March 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load tidyverse

```{r}
library(tidyverse)
```

## Create a data frame that just lists 

```{r}
data.frame(table(aw.df$h))
aw.df.2 <- aw.df %>% group_by(h) %>%tally()
```


## Create a density plot comparing abundance of Alewife by the hour of the day. 

```{r}
ggplot(aw.df,aes(h))+
  geom_density()+
 xlab("Hour of day")+
ggtitle("Amount of Alewife Captured Per Hour at Ahmerst and Missassguash Tidal Gates")


```
## Create a density plot comparing the two fishways *Ahmerst= cbcl, and Missaguash= mstdg) 

```{r}
target <- c("cbcl","mstgd")
temp_st <- filter(aw.df, site %in% target) 


```


```{r}
ggplot(temp_st,aes(h))+
  geom_density(mapping=aes(color=site))+
 xlab("Hour of day")+
ggtitle("Amount of Alewife Captured Per Hour:Ahmerst and Missassguash Tidal Gates")
```


## Install Gam package

```{r}
install.packages("gam")
install.packages("mgcv")
```


## Fit my data using a gam function instead of glm. 
## There no longer appears to be a significant difference between hour and abundance of Alewife caught. 


```{r}
library(mgcv)
pg1 <- gam(n ~ h,family = poisson, data = aw.df.2)

summary(pg1)
```



## Read in Tide Data 
## Change date and create an hour column. 

```{r}
library(readr)
tidal_cycles <- read_csv("tidal_cycles.csv", 
    col_types = cols(enter = col_date(format = "%m/%d/%Y")))
View(tidal_cycles)



```




## Seperate year month and day year into seperate columns in Tidal_cycles. 
## Create a column with just month and day. 

```{r}
library(lubridate)
tidal_cycles<- cbind(tidal_cycles, day=day(tidal_cycles$enter), month=month(tidal_cycles$enter), year=year(tidal_cycles$enter))



```

## Create a column that just includes month and day in tidal_cycles2

```{r}

tidal_cycles2 <- tidal_cycles %>%
  mutate(enter=make_date(!year,month,day))


```

## Graph the height of the tidal cycles by May and then June create two seperate data files. 

```{r}
tide_may <- filter(tidal_cycles2,month==5)
tide_june <- filter(tidal_cycles2,month==6)
```


```{r}
ggplot(data=tide_may)+
  geom_line(mapping=aes(x=h,y=Height))+
  facet_wrap(~day)+
  ylab("Height(m)")+
  xlab("Hour")+
  ggtitle("Change in Tidal Amplitude for May 2017")
```

## Graph the tidal amplitudes for every day in June. 

```{r}
ggplot(data=tide_june)+
  geom_line(mapping=aes(x=h,y=Height))+
  facet_wrap(~day)+
  ylab("Height(m)")+
  xlab("Hour")+
  ggtitle("Change in Tidal Amplitude for June 2017")
```

## Graph the height of the tide by the month and the day. 

```{r}
data_1_day <- tidal_cycles<- filter(tidal_cycles, enter=="2017-05-10")


ggplot(data=data_1_day)+
  geom_line(mapping=aes(x=h,y=Height))+
  ylab("Height(m)")+
  xlab("Hour")+
  ggtitle("Change in Tidal Amplitude for May 10 2017")
```

## Create a count of Alewife for each day by hour.
## change the enter in tidal_cycles to a character vector. 
## Sort it by the date(enter), left join aw.df.4 and tidal_cycles2 by hour and enter. 

```{r}
data.frame(table(aw.df$h,aw.df$enter))
aw.df.4 <- aw.df %>% group_by(h,enter) %>%tally()
aw.df.4 %>% arrange(enter)




```

## Change enter in tidal_cycles to a character vector so I can left join aw.df.4

```{r}
tidal_cycles2 <- tidal_cycles %>%
  mutate(enter=make_date(!year,month,day))

tidal_cycles2$enter <- as.character(tidal_cycles2$enter)
```

## Create columns for day, month, year in aw.df.4
## Change enter column to just month and day. 

```{r}
aw.df.4<- cbind(aw.df.4, day=day(aw.df.4$enter), month=month(aw.df.4$enter), year=year(aw.df.4$enter))

aw.df.4 <- aw.df.4 %>%
  mutate(enter=make_date(!year,month,day))


```



## Left Join tidal_cycles by enter and h to aw.df.4

```{r}
tides_alewife <- aw.df.4 %>% left_join(tidal_cycles2, by = "h","enter")


```


## Graph alewife abundance per hour with tidal height. 

```{r}

```







## Fit a poission model that has tidal height as a covariate. ## I am not sure how to do this. 

```{r}
 pg2 <-gam(n~h+Height,family=poisson,data=tides_alewife)
summary(pg2)


```


## Creat bins of tidal heights to graph. 

```{r}
n_height <- tides_alewife[,c("Height","n")]
```



```{r}

breaks <- c(0,1.5,3,4.5,6,7.5,9) # set up boundaries for bins
labels<- c("<1.5","1.5-3","3-4.5","4.5-6","6-7.5","7.5-9") # specify bin labels
bins <- cut(n_height$Height,breaks,include.lowest = T,right=FALSE,labels=labels)

summary(bins)

```

## Plot the bins
```{r}
plot(bins, main="Frequency", ylab= "Number of Alewife", xlab="Tidal Heights (m)",ggtitle="Frequency of Alewife versus Tidal Height between May 5th-June 12th 2017",col="blue")
```

## Graph the bins

```{r}
ggplot(data=bins)+
  geom_freqpoly()
```


## Graph abundance by tidal height. 

```{r}
ggplot(data=tides_alewife)+
  geom_point(aes(x=Height,y=n))
```

